A High-Order Discontinuous Galerkin Solver with Dynamic Adaptive Mesh Refinement to Simulate Cloud Formation Processes

### 2019-09-06 writing/slides/ppam.tex
Update slides.

\begin{frame}
  \frametitle{Two Bubbles: Simulation}
  %  \begin{figure}[h]
  %   \centering
  %   \ifratio{43}{
  %   \movie[width=0.9\textwidth, height=0.6\textwidth, autostart,, loop, poster]{}{two_bubbles_cartesian.ogv}
  % } {
  %   \movie[width=0.7\textwidth, height=0.466666\textwidth, autostart,, loop, poster]{}{two_bubbles_cartesian.ogv}
  % }
  %   \caption{Two Bubbles scenario}
  % \end{figure}
  Here: AMR simulation of two bubbles scenario (video)
\end{frame}

@@

\textbf{Discontinuous Galerkin} (\textsc{DG}) divides domain into disjoint elements, approximates solutions by piecewise-polynomials.
Elements are connected by solving the Riemann problem.

Implemented in the \textsc{PDE}-Framework \textbf{ExaHyPE}.
\end{frame}

@@

With $\Qrho$ density of fluid, $\Qj$ velocity density and $\QE$ energy density, $\bm{k}$ unit vector in z-direction.

@@

% \begin{frame}{The Navier-Stokes Equations}
%   TODO PDE Conservative form(?)
%   Vector of conserved quantities:
% \begin{equation}
%   \label{eq:conserved-variables}
%  \Q = \left( \Qrho, \Qj, \QE \right) 
% \end{equation}
% With $\Qrho$ density of fluid, $\Qj$ velocity density and $\QE$ energy density

% Flux:
% \begin{equation}
%   F(\Q, \textcolor{orange}{\gradQ}) = 
%   \begin{pmatrix}
%     \Qj \\
%     \Qv  \otimes \Qj + \bm{I} \pressure + \textcolor{orange}{\stressT (\Q, \gradQ)}  \\
%     \Qv \cdot \left(\bm{I} \QE + \bm{I} \pressure + \textcolor{orange}{\stressT (\Q, \gradQ)} \right) -
%     \textcolor{orange}{\kappa \nabla T}
%   \end{pmatrix}
% \end{equation}

% Pressure with gravitational term $\pressure(\Q, z)$,
% stress tensor $\textcolor{orange}{\stressT}$, heat diffusion $\textcolor{orange}{\kappa \nabla T}$ with temperature $T$.
% \end{frame}

\begin{frame}{Problem: Not hyperbolic}
  ExahyPE (thus far) solves equations of the form (e.g.\ Euler):
  \begin{equation}
  \label{eq:conservation-law}
  \mathbf{\textcolor{gray}{P}}\pdv{\Q}{t} + \divergence{\flux(\Q)}
 +  \textcolor{gray}{\sum_{i=1}^d \mathbf{B}_i (\mathbf{Q}) \pdv{\Q}{x_i}}
  = \source(\bm{x}, t, \Q)
\end{equation}

@@

\textbf{Solution}:
Modify numerical flux (Riemann solver), time step size (\textsc{cfl}-Condition) and boundary conditions to \textbf{allow diffusive terms}\footfullcite{dumbser2010arbitrary,gassner2008discontinuous}.

@@

\begin{frame}{Adaptive Mesh Refinement: Outliers}
  \begin{block}{Chebyshev's inequality}
\begin{equation}
  \label{eq:chebychev}
  \mathbb{P}(\vert X - \mu \vert \geq c \sigma) \leq \frac{1}{c^2}
\end{equation}
Mean $\mu$, standard deviation $\sigma$, constant $c$

@@

\begin{equation}
  \label{eq:refinement-criterion}
  \operatorname{evaluate-refinement}(\Q, \mu, \sigma) =
  \begin{cases}
    \text{refine} & \text{if } \tv(\Q) \geq \mu + T_\text{refine} \sigma \\
    \text{delete} & \text{if } \tv(\Q) < \mu + T_\text{delete} \sigma \\
    \text{keep} & \text{otherwise}
    \end{cases}
\end{equation}
Computation of mean $\mu$, standard deviation $\sigma$ with \textbf{stable}, pairwise reduction\footfullcite{chan1982updating}.
\end{block}
\end{frame}

@@

      $T_\text{refine} = 2.5$ and $T_\text{delete} = -0.5$.
    \item[Reference] $81 \times 81 = 6561$ cells.
    \end{description}
  \end{block}
  Both polynomial order 5. Viscosity $\mu = 0.01$. (?)

  AMR grid: 1953 cells.

  Relative $L_2$-error between AMR and fully refined: \num{6.69e-6}.
\end{frame}

\begin{frame}
  \frametitle{AMR vs.\ Fully Refined Grid: Grid}

  Here: Video/Figure of AMR grid
  %  \begin{figure}[h]
  %   \centering
  %   \ifratio{43}{
  %   \movie[width=0.9\textwidth, height=0.6\textwidth, autostart,, loop, poster]{}{two_bubbles_legendre.ogv}
  % }{
  %   \movie[width=0.7\textwidth, height=0.46666\textwidth, autostart,, loop, poster]{}{two_bubbles_legendre.ogv}
  % }
  %   \caption{Two bubbles scenario, with \amr{}}
  % \end{figure}
\end{frame}

\begin{frame}
  \frametitle{AMR vs.\ Fully Refined Grid: Error}

  Here: Figure AMR error on grid
%   \begin{figure}[H]
%     \centering
%     \ifratio{43}{
% %\includegraphics[width=0.9\textwidth]{thesis_two_bubbles_amr_error}
%   } {
% %\includegraphics[width=0.6\textwidth]{thesis_two_bubbles_amr_error}
%   }
% \caption{Potential temperature of fully refined solution minus \amr{} solution.}
%   \end{figure}
\end{frame}

\begin{frame}{3D Cosine Bubble}
  TODO
\end{frame}

\begin{frame}{MUSCL-Hancock Scheme~\footfullcite{vanLeer1979towards}}
  Finite Volume scheme: store only \textbf{cell averages}

@@

  \textbf{Second order} in time and space

  \textbf{Stabilized} with (minmod) slope limiter (TODO)
\end{frame}

\begin{frame}
  \frametitle{Two Bubbles: Finite Volume}
 TODO  
\end{frame}

\begin{frame}{Summary}
  \begin{block}{Results}
    \begin{itemize}
    \item \textbf{ADER-DG} can be used to simulate Navier-Stokes equations.
    \item Total variation \textbf{measures edges} of numerical solution.
    \item Chebyshev's criterion finds \textbf{interesting cells}.
    \item Combination of both \textbf{accurately tracks} cloud.
    \item AMR solution is close to fully refined solution but \textbf{faster}.
    \item \textbf{Artifical viscosity} of finite volume has a similar effect to \textbf{physical viscosity} of Navier-Stokes.  
    \end{itemize}
  % \begin{itemize}
  % \item \aderdg{} \textbf{converges} for Navier Stokes equations
  % \item \aderdg{} shows \textbf{good results} for large set of scenarios (many not shown here!)
  % \item \amr{}-criterion is \textbf{faster} and tracks cloud \textbf{correctly}.
  % \item (Limited) \aderdg{} works for reactive Euler \textit{\&} Navier-Stokes
  % \item All \textbf{extensions} to engine (diffusive fluxes, global observables) useful for \textbf{other users} of \exahype{}.
  % \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Acknowledgements}
This project has received funding from the European Union's Horizon 2020 research and
innovation programme under grant agreement No 823844 (\textbf{ChEESE}) and 671698 (\textbf{ExaHyPE}).

TODO: Grants correct?

\begin{center}
\includegraphics[width=0.7\textwidth]{logos}
\end{center}
\end{frame}
\end{document}

>>>>

\begin{frame}
  \frametitle{Two Bubbles: Simulation}
  \begin{center}
   \ifratio{43}{
%   \movie[width=0.9\textwidth, height=0.466666\textwidth, autostart,, loop, poster]{}{ppam_two_bubbles_cart.ogv}
 } {
   % todo: compute aspect ratio
   \movie[width=0.5\textwidth, height=0.31891168599\textwidth,autostart,, loop, poster]{}{ppam_two_bubbles_cart.ogv}
 }
  \end{center}
\end{frame}

@@

\textbf{Discontinuous Galerkin} (DG) divides domain into disjoint elements, approximates solutions by piecewise-polynomials.
Elements are connected by solving the Riemann problem.

\textbf{ADER}-Approach uses space-time polynomials for time integration instead of Runge-Kutta procedures.
\end{frame}

@@

With $\Qrho$ density of fluid, $\Qj$ velocity density, $\QE$ energy density, and $\bm{k}$ unit vector in z-direction.

@@

\begin{frame}{Problem: Not hyperbolic}
  ExaHyPE (thus far) solves equations of the form (e.g.\ Euler):
  \begin{equation}
  \label{eq:conservation-law}
  \mathbf{\textcolor{gray}{P}}\pdv{\Q}{t} + \divergence{\flux(\Q)}
 +  \textcolor{gray}{\sum_{i=1}^d \mathbf{B}_i (\mathbf{Q}) \pdv{\Q}{x_i}}
  = \source(\bm{x}, t, \Q)
\end{equation}

@@

\textbf{Solution}:
Modify numerical flux (Riemann solver), time step size (cfl-Condition) and boundary conditions to \textbf{allow diffusive terms}\footfullcite{dumbser2010arbitrary,gassner2008discontinuous}.

@@

\begin{frame}{Adaptive Mesh Refinement: Feature Detection}
  \begin{block}{Chebyshev's inequality}
\begin{equation}
  \label{eq:chebychev}
  \mathbb{P}(\vert X - \mu \vert \geq c \sigma) \leq \nicefrac{1}{c^2}
\end{equation}
Mean $\mu$, standard deviation $\sigma$, constant $c$

@@

\begin{equation}
  \label{eq:refinement-criterion}
  \operatorname{evaluate-refinement}(\Q, \mu, \sigma) =
  \begin{cases}
    \text{refine} & \text{if } \tv\left(\bm{f}(\Q)\right) \geq \mu + T_\text{refine} \sigma \\
    \text{delete} & \text{if } \tv\left(\bm{f}(\Q)\right) < \mu + T_\text{delete} \sigma \\
    \text{keep} & \text{otherwise}
    \end{cases}
\end{equation}
Choose $T_\text{refine}$ and $T_\text{delete}$ according to cost-quality trade-off.

Computation of mean $\mu$, standard deviation $\sigma$ with \textbf{stable}, pairwise reduction\footfullcite{chan1982updating}.
\end{block}
\end{frame}

@@

      $T_\text{refine} = 2.5$ and $T_\text{delete} = -0.5$.
    \item[Reference] $81 \times 81 = 6561$ cells.
    \end{description}
  \end{block}
  Both polynomial order 6. Viscosity $\mu = 0.01$.

  AMR grid: 1953 cells. Less than \textbf{30\%} of full grid!

  Relative $L_2$-error between AMR and fully refined: \num{2.6e-6}.
\end{frame}

\begin{frame}
  \frametitle{AMR vs.\ Fully Refined Grid: Grid}
  %  \begin{figure}[h]
  %   \centering
  %   \ifratio{43}{
  %   \movie[width=0.9\textwidth, height=0.6\textwidth, autostart,, loop, poster]{}{two_bubbles_legendre.ogv}
  % }{
  \begin{center}
   \movie[width=0.5\textwidth, height=0.318911685990\textwidth, autostart,, loop, poster]{}{ppam_two_bubbles_leg.ogv}
  \end{center}
  % }
  %   \caption{Two bubbles scenario, with \amr{}}
  % \end{figure}
\end{frame}

\begin{frame}
  \frametitle{AMR vs.\ Fully Refined Grid: Error}
  \begin{figure}[H]
     \centering
%     \ifratio{43}{
% %\includegraphics[width=0.9\textwidth]{thesis_two_bubbles_amr_error}
%   } {
\includegraphics[width=0.55\textwidth]{ppam_two_bubbles_amr_error}
%   }
 \caption*{Potential temperature of fully refined solution minus AMR solution.}
   \end{figure}
\end{frame}

\begin{frame}{3D Cosine Bubble\footfullcite{kelly2012continuous}}
  \begin{columns}
    \begin{column}{0.6\textwidth}
  \begin{center}
    \includegraphics[width=0.85\textwidth]{ppam_cosine_bubble_3d}
  \end{center}
    \end{column}
    \begin{column}{0.4\textwidth}
      \begin{itemize}
      \item Polynomial Order 3
      \item $25^3$ cells
      \item Time: \SI{400}{\s}
      \item Viscosity: $\mu = 0.05$
      \end{itemize}
    \end{column}

  \end{columns}
\end{frame}

\begin{frame}{MUSCL-Hancock Scheme}
  Finite Volume scheme: store only \textbf{cell averages}

@@

  \textbf{Second order} in time and space\footfullcite{vanLeer1979towards}

  \textbf{Stabilized} with (Van Albada~\footfullcite{van1997comparative}) slope limiter

  Very stable, larger \textbf{numerical viscosity}
\end{frame}

\begin{frame}
  \frametitle{Two Bubbles: Finite Volume}
  \begin{center}
    \includegraphics[width=0.6\textwidth]{paper_two_bubbles_fv}
  \end{center}
\end{frame}

\begin{frame}{Summary}
  \begin{block}{Results}
    \begin{itemize}
    \item \textbf{ADER-DG} can be used to simulate Navier-Stokes equations.
    \item Total variation \textbf{measures edges} of numerical solution.
    \item Chebyshev's criterion finds \textbf{interesting cells}.
    \item Combination of both \textbf{accurately tracks} cloud.
    \item AMR solution is close to fully refined solution but needs \textbf{fewer} cells.
    \item \textbf{Artificial viscosity} of finite volume has a similar effect as \textbf{physical viscosity} of Navier-Stokes.  
    \end{itemize}
  % \begin{itemize}
  % \item \aderdg{} \textbf{converges} for Navier Stokes equations
  % \item \aderdg{} shows \textbf{good results} for large set of scenarios (many not shown here!)
  % \item \amr{}-criterion is \textbf{faster} and tracks cloud \textbf{correctly}.
  % \item (Limited) \aderdg{} works for reactive Euler \textit{\&} Navier-Stokes
  % \item All \textbf{extensions} to engine (diffusive fluxes, global observables) useful for \textbf{other users} of \exahype{}.
  % \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Acknowledgments}
This project has received funding from the European Union's Horizon 2020 research and
innovation programme under grant agreement No 823844 (\textbf{ChEESE}) and 671698 (\textbf{ExaHyPE}).

\begin{center}
\includegraphics[width=0.7\textwidth]{logos}
\end{center}
\end{frame}

\begin{frame}
  \frametitle{Shared-Memory Scaling (from\footfullcite{reinarz2019exahype})}

  \begin{center}
  \includegraphics[width=0.4\textwidth]{speedup_tbb}
  \end{center}

\end{frame}
\end{document}
